---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# NormalRetina

<!-- badges: start -->
[![R-CMD-check](https://github.com/UofUEpiBio/egpkg/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/UofUEpiBio/egpkg/actions/workflows/R-CMD-check.yaml)
<!-- badges: end -->

The goal of NormalRetina is to predict normative retina sensitivity, providing tools for data processing, modeling, prediction and visualization. It is particularly useful for age-adjusted normative data for retinal sensitivity.

## Installation

You can install the development version of NormalRetina from [GitHub](https://github.com/LydiaYW/NormalRetina) with:

``` r
# install.packages("devtools")
devtools::install_github("LydiaYW/NormalRetina")
```



## Features

- **Data Preprocessing** *SensForFit*: A function to clean and prepare retinal sensitivity data for analysis, with variables: participant age, eye, eccentricity and angular distance from the fovea, and spatial coordinates (x,y). Also patient-wise k-fold split for model training.
- **Virtual Datasets**: Includes two example datasets representing normative retinal sensitivity for reference patients. First is *ref77*, a virtrual patient of age 77, for spatial interpolation and visualization. The other is *refMes*, a virtrual population with 100 healthy participants' Mesopic sensitivity results.
- **Prediction Models**:
  - Linear Mixed Model (LMM)
  - Bayesian Quantile Regression (BQR)
  - Random Forest (RF)
  - These models predict age-adjusted normative data based on input features. Also performance measures for model comparison.
- **Spatial Interpolation**: A method for interpolating retinal sensitivity for patients of a specific age.

---


## Example

This is a basic example which shows you how to solve a common problem:

```{r example}
library(NormalRetina)
## basic example code
```
